The Community Land Model, version 5.1 \\ One-at-a-time Parameter Perturbation Ensemble

### 2023-08-02 ms/ms.tex
add refs

\subsection{Experimental Design}
Identification of the complete parameter set of a land surface model is in itself a non-trivial exercise, as in practice, many empirical constants previously considered as `internal' are hard-coded and not amenable to systematic perturbation. For the purposes of this activity, we identified what we consider the full set of 188 CLM-BGC parameters, and enabled their modification via the model input files.

@@

In this project, we identified and perturbed 193 CLM parameters to create a large one-at-a-time PPE. This ensemble is useful for understanding parametric controls on CLM processes and has already been leveraged as a starting point for parameter optimization efforts. The software and analysis tools generated to enable this experiment will reduce the burden of generating future PPEs.

There were several barriers to perturbing the full set of CLM5-BGC parameters. Firstly, many parameters had not been officially identified as such. In such cases, we identified hard-coded values, established an appropriate parameter name, and extracted that parameter to the CLM parameter file for easier manipulation. In the end, we perturbed 193 parameters across a variety of CLM processes (Figure \ref{fig:params}). The second challenge involved defining a perturbation range for each parameter. We solicited expert judgment to set a minimum and maximum reasonable value for each parameter. In some cases, literature values were explicitly used (e.g. medlynslope), but the most common range was $\pm20\%$.

The third challenge involved managing computational cost.  Quantifying parameter effects is a necessary prerequisite to automated calibration and uncertainty quantification. Because the parameter space of CLM is quite large and the model response potentially non-linear, large ensembles are required to adequately resolve response surfaces. With standard CLM configurations, such ensembles would far exceed our computing allocation. As such, the computational cost of inferring parameter effects is a major constraint. We were able to speed up the model approximately 500x by strategically reducing the number of model grid cells (Figure \ref{fig:sg}) and by leveraging a linearized spin-up solver (see Sections \ref{sect:mcn} and \ref{sect:sg} for details). Designing faster model configurations and more efficient sampling strategies will greatly aid in model calibration and uncertainty quantification.

For this experiment we opted for a one-at-a-time perturbation strategy, testing a minimum and maximum value for each of 193 parameters. We found that parameter effects could be quite large, in some cases exceeding the effects of our various forcing scenarios (Figure \ref{fig:ranges}). That being said, the majority of parameter perturbations had very small effects, such that a majority of simulations were clustered around the default simulation. In the case of canopy evaporation, more than 95\% of the ensemble variance could be explained by the ten most influential parameters (Figure \ref{fig:variance}). This indicates that there may be very tractable parameter estimation sub-problems if global calibration of 193 parameters proves challenging.

@@

A one-at-a-time PPE cannot capture parameter interactions, and our min/max sampling protocol precludes diagnosing non-linearities. As such, this dataset will be insufficient for most calibration activities or for estimating overall parametric uncertainty. The advantage of our dataset is that we can diagnose parameter effects without the uncertainty contributed by an  emulator or regression model. As such, it is easy to diagnose which parameters are most influential on a given process, which has already been leveraged for parameter filtering. We have published a large set of ensemble diagnostic plots online, which serves as a valuable enhancement to our model technical documentation. Now in addition to seeing the definition of a given parameter, and the relevant equations, a model user can easily investigate the magnitude and spatial patterns of its effects (e.g. Figure \ref{fig:panel}).

We have already begun a follow-on experiment that perturbs a subset of parameters with the Latin hypercube sampling strategy to work towards calibrating leaf area index. By investing in the infrastructure underlying this first experiment, all of our subsequent experiments have required much less time and effort. We expect to continue to extend our efforts in this domain, as well as repeat these types of experiments with subsequent model releases.

>>>>

\subsection{Experimental Design}
Identification of the complete parameter set of a land surface model is in itself a non-trivial exercise, as in practice, many empirical constants previously considered as `internal' are hard-coded and not amenable to systematic perturbation. For the purposes of this activity, we identified what we consider a very broad set of 193 CLM-BGC parameters, and enabled their modification via the model input files. We opted not to perturb crop parameters within this ensemble, as our focus is on unmanaged vegetation.

@@

In this project, we identified and perturbed 193 CLM parameters to create a large one-at-a-time PPE. This ensemble is useful for understanding parametric controls on CLM processes. The software and analysis tools generated to enable this experiment will greatly reduce the burden of generating future PPEs and parameter sensitivity experiments.

There were several barriers to perturbing the full set of CLM5-BGC parameters. Firstly, many parameters had not been officially identified as such. In such cases, we identified hard-coded values, established an appropriate parameter name, and extracted that parameter to the CLM parameter file for easier manipulation. In this experiment we perturbed 193 parameters across a variety of CLM processes (Figure \ref{fig:params}). This is not the full set of CLM parameters, as some processes were not included, such as crops. In carrying out this experiment, we have likewise realized that the epistemology of climate model `parameters' is highly nuanced, and that defining the parameters within a given model structure can be somewhat subjective, such that it would be very difficult to collate a comprehensive or definitive set of parameters for a model like CLM.

The second challenge involved defining a perturbation range for each parameter. We solicited expert judgment to set a minimum and maximum reasonable value for each parameter. In some cases, literature values were explicitly used (e.g. medlynslope, \citeA{lin2015}), but the most common range was $\pm20\%$. It is exceedingly difficult to set parameter ranges that sample comparable probability density, even in a univariate experiment, like this one. The number of parameters is very large, many of which lack sufficient empirical backing for robust range evaluation. Even with parameters that have an empirical basis, they may behave differently at the coarse climate modeling spatial resolution, complicating the process of uncertainty quantification. As such, we cannot claim that the parameters are equivalently sampled, whereby parameter effects and parameter rankings will be subject to sampling asymmetries.

The third challenge involved managing computational cost.  Quantifying parameter effects is a necessary prerequisite to automated calibration and uncertainty quantification. Because the parameter space of CLM is quite large and the model response potentially non-linear, large ensembles are required to adequately resolve response surfaces. With standard CLM configurations, such ensembles would far exceed our computing allocation. As such, the computational cost of inferring parameter effects is a major constraint. We were able to speed up the model approximately 500x by strategically reducing the number of model grid cells (Figure \ref{fig:sg}) and by leveraging a linearized spin-up solver (see Sections \ref{sect:mcn} and \ref{sect:sg} for details). As long as computational resources remain constrained, designing faster model configurations and more efficient sampling strategies will be important for effective model calibration and uncertainty quantification.

For this experiment we opted for a one-at-a-time perturbation strategy, testing a minimum and maximum value for each parameter. We found that parameter effects could be quite large, in some cases exceeding the effects of our various forcing scenarios (Figure \ref{fig:ranges}). That being said, the majority of parameter perturbations had very small effects, such that a majority of simulations were clustered around the default simulation. In the case of canopy evaporation, more than 95\% of the ensemble variance could be explained by the ten most influential parameters (Figure \ref{fig:variance}). This indicates that there may be very tractable parameter estimation sub-problems if global calibration of 193 parameters proves unattainable.

@@

A one-at-a-time PPE cannot capture parameter interactions, and our min/max sampling protocol precludes diagnosing non-linearities. As such, this dataset will be insufficient for most calibration activities or for estimating overall parametric uncertainty. The advantage of our dataset is that we can diagnose parameter effects without the uncertainty contributed by an emulator or regression model. As such, it is easy to diagnose which parameters are most influential on a given process. We have published a large set of ensemble diagnostic plots online, which serves as a valuable enhancement to our model technical documentation. Now in addition to seeing the definition of a given parameter, and the relevant equations, a model user can easily investigate the magnitude and spatial patterns of its effects (e.g. Figure \ref{fig:panel}).

Several spin-off projects have leveraged the work presented here, including some that have already reached publication status \cite{cheng2023,yan2023a,yan2023b}. These projects utilized the output from this ensemble to filter for parameters that influence the relevant study domains and used the parameter ranges collated in Section zqz. Our project has accelerated parameter exploration work within our collaborator network by providing:

\begin{itemize}
\item Parameter ranges for nearly 200 parameters
\item Machine learning smart low resolution unstructured grid (sparsegrid)
\item Accelerated spinup procedure
\item Ensemble generation scripting toolchain
\item Parameter sensitivity diagnostics for nearly 200 parameters across 6 forcing scenarios
\end{itemize}

We have likewise begun a follow-on experiment that perturbs a subset of parameters with the Latin hypercube sampling strategy to work towards calibrating leaf area index. By investing in the infrastructure underlying this first experiment, all of our subsequent experiments have required much less time and effort. We expect to continue to extend our efforts in this domain, as well as repeat these types of experiments with subsequent model releases.
